### 2020-12-16 clean data scripts/clean_survey_data.Rmd

---
title: "clean_survey_data_from_CES"
author: "Qiyue Zhang"
date: "16/12/2020"
output: html_document
---
```{r setup, include=FALSE}
#set up libraries
knitr::opts_chunk$set(echo = TRUE)
library(haven)
library(tidyverse)
library(dplyr)
library(cesR)
library(labelled)
```

```{r}
# Grabbing ces2019_web dataset
ces2019_web <- read_dta("ces2019_web.dta")
```

```{r}
# Converting dataset to factor form
ces2019_web <- labelled::to_factor(ces2019_web)
```

```{r}
# Keep variables that we're interested
reduced_data <- 
  ces2019_web %>% 
  select(cps19_gender,
         cps19_education,
         cps19_employment,
         cps19_bornin_canada,
         cps19_province,
         cps19_votechoice,
         cps19_income_number,
         pes19_age
         )
```

```{r}
# rename some variables 
survey <- reduced_data %>% rename(
  sex = cps19_gender,
  votechoice = cps19_votechoice,
  edu = cps19_education,
  income = cps19_income_number,
  employed = cps19_employment,
  bornin_canada = cps19_bornin_canada,
  province = cps19_province,
  age = pes19_age
  )
```

```{r}
## Clean up##
#Rename gender variable to keep consistency with census data:
survey <- survey %>%
  mutate(sex = case_when(
    sex == "A woman" ~ "Female",
    sex == "A man" ~ "Male",
    TRUE ~ "NA"
  ))

#Simplify the voting choice variable by reducing the number of categories:
survey <- survey %>%
  mutate(votechoice = case_when(
    votechoice == "Liberal Party" ~ "Liberal Party",
    votechoice == "Conservative Party" ~ "Conservative Party",
    TRUE ~ "NA"
  ))

#Change education data in to binary 1/0 (college or not) variable:
survey <- survey %>%
  mutate(edu = case_when(
    edu == "Some elementary school" ~ 0L,
    edu == "No schooling" ~ 0L,
    edu == "Completed elementary school" ~ 0L,
    edu == "Don't know/ Prefer not to answer" ~ NA_integer_,
    edu == "Completed secondary/ high school" ~ 0L,
    edu == "Some secondary/ high school"~ 0L,
    TRUE ~ 1L
  ))

#Change income variable into income groups to keep consistency with census data:
survey <- survey %>% mutate(income = case_when(income >= 25000  & income <= 49999 ~ '$25,000 to $49,999',
                                             income >= 75000  & income <= 99999 ~ '$75,000 to $99,999',
                                             income >= 100000  & income <= 124999 ~ '$100,000 to $ 124,999',
                                             income >= 50000  & income <= 74999 ~ '$50,000 to $74,999',
                                             income < 25000 ~ 'Less than $25,000',
                                             income > 125000 ~ '$125,000 and more',
                                             TRUE ~ "NA"))

#Change employed data in to binary 1/0 (employed for pay or not) variable:
survey <- survey %>%
  mutate(employed = case_when(
    employed == "Don't know/ Prefer not to answer" ~ NA_integer_,
    employed == "Student" ~ 0L,
    employed == "Unemployed/ looking for work" ~ 0L,
    employed == "Self employed (with or without employees)" ~ 0L,
    employed == "Retired" ~ 0L,
    employed == "Disabled" ~ 0L,
    employed == "Caring for a family"~ 0L,
    employed == "Other (please specify)" ~ NA_integer_,
    TRUE ~ 1L
  ))

#Change born in canada data in to binary 1/0 (born in canada or not) variable:
survey <- survey %>%
  mutate(bornin_canada = case_when(
    bornin_canada == "Don't know/ Prefer not to say" ~ NA_integer_,
    bornin_canada == "No" ~ 0L,
    TRUE ~ 1L
  ))

#Simplify province data by reducing the number of categories to keep consistency with census data:
survey <- survey %>%
  mutate(province = case_when(
    province == "Northwest Territories" ~ "NA",
    province == "Nunavut" ~ "NA",
    province == "Yukon" ~ "NA",
    province == "Quebec" ~ "Quebec",
    province == "Manitoba" ~ "Manitoba",
    province == "Ontario" ~ "Ontario",
    province == "Alberta" ~ "Alberta",
    province == "Nova Scotia" ~ "Nova Scotia",
    province == "British Columbia" ~ "British Columbia",
    province == "Saskatchewan" ~ "Saskatchewan",
    province == "Prince Edward Island" ~ "Prince Edward Island",
    province == "New Brunswick" ~ "New Brunswick",
    province == "Newfoundland and Labrador" ~ "Newfoundland and Labrador"
  ))

#Change age variable into age groups with exception of under 18: 
survey <- survey %>% mutate(age = case_when(age >= 18  & age <= 29 ~ '18 to 29',
                                             age >= 30  & age <= 44 ~ '30 to 44',
                                             age >= 45  & age <= 59 ~ '45 to 59',
                                             age >= 60  & age <= 74 ~ '60 to 74',
                                             age > 75 ~ '75 and above',
                                             TRUE ~ "NA"))
```

```{r}
#Create a .csv file
write.csv(survey, "survey_cleaned.csv")
```