### 2020-12-17 clean data scripts/clean_census_data.Rmd
model and result to be fixed

---
title: "clean_census_data_from_GSS"
author: "Qiyue Zhang"
date: "16/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#workplace setup
library(haven)
library(tidyverse)
library(dplyr)
library(readr)
```

```{r}
#import gss data
gss <- read_csv("gss.csv")
```

```{r,warning=FALSE, message=FALSE}
# Keep variables that we're interested
reduced_dat <- 
  gss %>% 
  select(sex,
         education,
         worked_last_week,
         place_birth_canada,
         province,
         age
         )
```

```{r}
# rename some variables 
census <- reduced_dat %>% rename(
  edu = education,
  employed = worked_last_week,
  bornin_canada = place_birth_canada)
```

```{r}
## Clean up##
#Change education data in to binary 1/0 (college or not) variable:
census <- census %>%
  mutate(edu = case_when(
    edu == "High school diploma or a high school equivalency certificate" ~ 0L,
    edu == "Trade certificate or diploma" ~ 0L,
    edu == "Less than high school diploma or its equivalent" ~ 0L,
    edu == "NA" ~ NA_integer_,
    TRUE ~ 1L
  ))

#Change employed data in to binary 1/0 (employed for pay or not) variable:
census <- census %>%
  mutate(employed = case_when(
    employed == "No" ~ 0L,
    employed == "Don't know" ~ NA_integer_,
    employed == "NA" ~ NA_integer_,
    TRUE ~ 1L
  ))

#Change born in canada data in to binary 1/0 (born in canada or not) variable:
census <- census %>%
  mutate(bornin_canada = case_when(
    bornin_canada == "Born outside Canada" ~ 0L,
    bornin_canada == "NA" ~ NA_integer_,
    bornin_canada == "Don't know" ~ NA_integer_,
    TRUE ~ 1L
  ))

#Change age variable into age groups with exception of under 18: 
census <- census %>% mutate(age = case_when(age >= 18  & age <= 29 ~ '18 to 29',
                                             age >= 30  & age <= 44 ~ '30 to 44',
                                             age >= 45  & age <= 59 ~ '45 to 59',
                                             age >= 60  & age <= 74 ~ '60 to 74',
                                             age > 75 ~ '75 and above',
                                             TRUE ~ "NA"))
```

```{r}
#Create cell counts:
cell_counts <- census %>% group_by(province, sex, as.factor(age), employed, bornin_canada, edu) %>% count() %>% rename(age = "as.factor(age)")
#Create a .csv file
write.csv(cell_counts, "census_cleaned.csv")
```